---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Multiple Imputation by Super Learning (MISL)

[![Travis-CI Build Status](https://travis-ci.com/carpenitoThomas/misl.svg?token=u9TyfsxVjq6xxvc5h473&branch=master)](https://travis-ci.com/carpenitoThomas/misl)
[![Appveyor Build Status](https://ci.appveyor.com/api/projects/status/hagh8vidrdeacr7f?svg=true)](https://ci.appveyor.com/project/tlverse/sl3)
[![Coverage Status](https://img.shields.io/codecov/c/github/tlverse/sl3/master.svg)](https://codecov.io/github/tlverse/sl3?branch=master)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![License: GPL v3](https://img.shields.io/badge/License-GPL%20v3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1342293.svg)](https://doi.org/10.5281/zenodo.1342293)



The goal of MISL (Multiple Imputation by Super Learning) is to create multiply imputed datasets using the super learning framework. This package builds heavily off of the `sl3` and `mice` packages.

## Installation

You can install the released version of misl from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("misl")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("carpenitoThomas/misl")
```
## Example

Here's an example with the nhanes data in which we use `misl()` imputation and then pool the results:

```{r example}
library(misl)

misl_imp <- misl(nhanes, maxit = 2, m = 2, quiet = TRUE)

misl_modeling <- lapply(misl_imp, function(y){
  stats::lm(TotChol ~ Age + Weight + Height + Smoke100 + Education, data = y)
})

summary(mice::pool(misl_modeling))
```
